---
title: "Structured Population Models"
author: "Anoob Prakash"
date: "30 January 2019"
output: html_document
---

```{r}
load("COMPADRE_v.4.0.1.RData")
ls(compadre)
head(compadre$metadata)
```

#### Species: *Centaurea horrida*

This is an endagered species according to the IUCN Red List, with two extant population's in Sardinia with about total recorded population size < 12,000. These species posses limited ability to disperse as well as colonize, leading to their limited range. Eventhough it's population is at present stable, it's low population size and limited range puts it under risk of going extinction in case of some drastic changes in the surrounding environment. We would like to see if the species would go extinct in the near future or not. The genus is also of particular interest for the Keller Lab. 

```{r}

library(tidyverse)

load("./COMPADRE_v.4.0.1.RData")

grep(compadre$metadata$SpeciesAccepted, pattern = "Centaurea horrida") 
# six entries

#get matrixA from most recent entry 

mat1 <- compadre$mat[1496][[1]]$matA

```
this package makes it all much easier 
```{r}

#install.packages("demogR")
library(demogR)

```
first eigenvalue
```{r}
eigen.analysis(mat1)[1]

```
stable age distrbution
```{r}
eigen.analysis(mat1)[5]
```

elasticities
```{r}
eigen.analysis(mat1)[4]
```

sensitivities

```{r}
eigen.analysis(mat1)[3]
```
It looks like the survival rate of the oldest individuals (bottom right value) is the most sensitive parameter

attempt to plot this population through time:
```{r}
# fuction to find population matrix for given year (t)
modl <- function(L,I,t) {
  N <- L^t %*% I
  return(N)
}

# need an input matrix
# let's say we start with 3 individuals in each class
input <- rep(3,5)

# trying the function

modl(mat1,input,1)
```

```{r}
# now try to model this population over a longer time scale
# 30 years
tVals <- seq(1,30)

# starting with 90 individuals in each class
input <- rep(90,5)

# iterate model

out <- data_frame(t= tVals,N = tVals)

for (i in 1:length(tVals)) {
out[i,2] <- sum(modl(mat1,input,tVals[i]))
}

# plot population over time

ggplot(out, aes(x=t, y= N)) +
  geom_point() 



```

plot each age class? over time

```{r}
#t column needs to be each time repeated 5 times, N needs to be length tVals*5 and new column class needs to be 1-5 over and over again
# make the t column within the loop

size <- length(tVals)*5
out2 <- data_frame(t= 1:size, N = 1:size, class = rep(1:5,length(tVals)))
countr <- 0
for (i in 1:length(tVals)) {
  r1 <- countr + i
  r2 <- (countr + i)+4
out2[r1:r2 , 2] <- modl(mat1,input,tVals[i])
#still need to make t column
#out2[i:]
countr <- countr + 4
}
```

#what would happen if you raised a bunch of plants to the oldest age class in a greenhouse and then released them?
# what about the timing of the realease 